Kindly Bent to Free Us

### 2019-05-27 kindly/infer/infer.tex
quantify over kvars
semantics & syntax-driven typing
inconsistencies fixed
Get rid of one notion of spliting.
Simplify/factor the region/borrowing rules.
Use types in use-maps.
Use the (\E <= k) syntax everywhere.
Reordering.

\subsubsection{Constraint language}

\newcommand\A{\mathcal A}
\newcommand\SC{\mathcal S}

Let us note $\A$ our constraint system. The grammar of constraints is
given in  \cref{grammar:constraint}.
$\A$ is defined as the smallest cylindric term constraint system that
satisfies the axiom shown in \cref{rules:entail}.
We follows the traditional HM(X) formulation
with conjunctions, projections and type inequalities.
The new element specific to our approach are kind inequalities.
Entailment is noted $\entail{C}{D}$, where $D$ is a consequence of the
constraints $C$.
We say that $C$ and $D$ are equivalent, noted $C \equivC D$,
when $\entail{C}{D}$ and $\entail{D}{C}$.
\TODO{Give the cylindric properties ?}

\begin{figure}[tp]
  \centering
  \begin{align*}
    C ::=&\ \Cleq{\tau_1}{\tau_2}\ |\ \Cleq{k_1}{k_2}\ |\ C_1 \Cand C_2\ |\ \Cproj{\alpha}{C}
  \end{align*}
  \caption{The constraint language}
  \label{grammar:constraint}
\end{figure}

\begin{figure*}[tp]
  \input{infer/lattice}
  \caption{Lattice inequalities -- $k \lk_\Lat k'$}
  \input{infer/entails}
  \caption{Base entailment rules -- $\entail{C}{D}$ }
  \label{rules:entail}
\end{figure*}

We note $\SC$ the set of solved forms
which can be used inside type and kind schemes.
We define $\SC$ as $\A$ quotiented by the relation $\equivC$.
%
We consider the existence of a function $\operatorname{normalize}$ which takes
a constraint in $\A$ and a substitution $\psi$ and returns a constraint
in solved form $C' \in \SC$,
and an updated substitution. We detail the implementation
of the normalization function in \cref{sec:normalize}

% $\mathcal S$ is composed only of kind
% inequalities \emph{over variables}. For convenience, if $C\in\mathcal S$, we
% note $C$ as a list of kind inequalities: $\Cleq{\kvar_i}{\kvar_{i'}}^n$.
% \TODO{Extend the properties of solved forms}

@@

\begin{figure}[tp]\centering
\begin{mathpar}
  % \Sv\cap\Sv'
  % &= \left\{ x \to l,l'
  %   \mid \Sone{x}{l}\in\Sv \text{ and } \Sone{x}{l'}\in\Sv' \right\} \\
  \infer
  { \forall x \in \Dom{\Sv_1}\cup\Dom{\Sv_2}.\quad
    \tsplit{C_x}{\Sv(x)}{\Sv_1(x)}{\Sv_2(x)}
  } 
  { \tsplit{(\bigwedge C_x)}{\Sv}{\Sv_1}{\Sv_2} }
  % &= \Sv\cap\Sv' \sqcup
  %   (\Sv\Sdel{x' \mid x'\in\Sv'}) \sqcup
  %   (\Sv'\Sdel{x \mid x\in\Sv})
  %   &\text{where $\sqcup$ is the disjoint union}\\
  \and
  \inferrule{}
  { \Cleq{\Sv}{k_0}
    = \bigwedge_{\Sone{x}{l_x}\in\Sv} \bigwedge_{k\in l_x} \Cleq{k}{k_0}
  }
  \and
  \inferrule
  { \forall \bshadow{x} \in \Sv, b'\in\Sv'.\quad
    \tregion{C_x}{y}{\bshadow{x}}{b'} }
  { \tregion{(\bigwedge C_x)}{y}{\Sv}{\Sv'} }
  \and
  % \inferrule
  % { \text{if } \operatorname{lenght}(\Sv(x))  = \Cempty{} \wedge  }
  { \Weaken_{(x,k)}(\Sv) =
    \begin{dcases}
      \Ctrue& \operatorname{length}(\Sv(x)) = 1\\
      {k}\lk\kaff_\infty& \text{ otherwise}\\
    \end{dcases}

@@

\begin{figure}[ht]
  \centering
  \input{infer/kinds}
  \caption{Kind inference algorithm -- $\inferK{(C,\unif)}{\E}{\tau}{k}$}
  \label{rules:kinding}
\end{figure}

\begin{figure}[ht]
  \centering
  \input{infer/sd-kinds}
  \caption{Syntax-directed kind inference -- $\inferSK{C}{\E}{\tau}{k}$}
  \label{rules:sd-kinding}
\end{figure}

@@

\begin{figure*}[h]
  \input{infer/typing}
  \caption{Type inference algorithm -- $\inferW{\Sigma}{(C,\psi)}{\bf{\E}}{\bf{e}}{\tau}$ }
  \label{rules:typing}
\end{figure*}

\begin{figure*}[ht]
  \input{infer/sd-typing}
  \caption{Syntax-directed typing rules}
  \label{fig:syntax-directed-typing}
\end{figure*}
\subsection{Normal forms and generalization}
\label{sec:normalize}

>>>>

@@

\begin{figure}[tp]\centering
\begin{mathpar}
  % \Sv\cap\Sv'
  % &= \left\{ x \to l,l'
  %   \mid \Sone{x}{l}\in\Sv \text{ and } \Sone{x}{l'}\in\Sv' \right\} \\
  % \infer
  % { \forall x \in \Dom{\Sv_1}\cup\Dom{\Sv_2}.\quad
  %   \bsplit{C_x}{\Sv(x)}{\Sv_1(x)}{\Sv_2(x)}
  % } 
  % { \bsplit{(\bigwedge C_x)}{\Sv}{\Sv_1}{\Sv_2} }
  % &= \Sv\cap\Sv' \sqcup
  %   (\Sv\Sdel{x' \mid x'\in\Sv'}) \sqcup
  %   (\Sv'\Sdel{x \mid x\in\Sv})
  %   &\text{where $\sqcup$ is the disjoint union}\\
  % \and
  \inferrule{}
  { \Cleq{\Sv}{k}
    = \bigwedge_{\Sone{x}{\schm}\in\Sv} \Cleq{\schm}{k}
  }
  \and
  % \inferrule
  % { \forall \bshadow{x} \in \Sv, b'\in\Sv'.\quad
  %   \tregion{C_x}{y}{\bshadow{x}}{b'} }
  % { \bregion{(\bigwedge C_x)}{y}{\Sv}{\Sv'} }
  % \and
  % \inferrule
  % { \text{if } \operatorname{lenght}(\Sv(x))  = \Cempty{} \wedge  }
  { \Weaken_{(x,k)}(\Sv) =
    \begin{dcases}
      \Ctrue& \operatorname{length}(\Sv(x)) = 1\\
      {k}\lk\kaff_\infty& \text{ otherwise}\\
    \end{dcases}

@@

\begin{figure}[ht]
  \centering
  \input{infer/kinds}
  \caption{Kind inference algorithm -- $\inferK{(C,\unif)}{\E}{\tau}{(\Multi{k}) k}$}
  \label{rules:kinding}
\end{figure}

@@

\begin{figure*}[htp]
  \input{infer/typing}
  \caption{Type inference algorithm -- $\inferW{\Sigma}{(C,\psi)}{\bf{\E}}{\bf{e}}{\tau}$ }
  \label{rules:typing}
\end{figure*}

\subsection{Normal forms and generalization}
\label{sec:normalize}

### 2019-05-27 kindly/lang.tex
minor tweaks for readability
Inference rules for regions.
semantics
Add typing for pairs and complete the kinds/lattices.
Smooth over the borrow-related bits.
update
Get rid of one notion of spliting.
Simplify/factor the region/borrowing rules.
Reordering.

The syntax of \lang is presented in \cref{grammar} and follows
traditional ML languages with let bindings.
Data constructors are introduced by $\introK{K}{e}$ and eliminated
with by $\elimK{K}{e}$.
Types are composed of type variable $\tvar$, parameterized type constructors
$\T{t}$ and arrow types annotated with a kind, noted $\tau \tarr{k} \tau$.
Kinds can be either a kind variable $\kvar$, affine $\kaff$ or unrestricted $kun$.
Type schemes and kind schemes are guarded by a set of constraints.

@@

\begin{figure*}[h]
  \centering
  \input{grammar}
  \caption{Syntax}
  \label{grammar}
\end{figure*}

\subsection{Semantics}

\TODO{}
% \begin{figure*}[h]
%   \input{semantics}
%   \caption{Reduction rules -- $\ered{I}{E}{e}{v}$ }
% \end{figure*}

\subsection{Typing}

\TODO{Add system-F like type rules}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:

>>>>

The syntax of \lang is presented in \cref{grammar} and follows
traditional ML languages with let bindings.
Data constructors are introduced by $\introK{K}{e}$ and eliminated
with by $\elimK{K}{e}$.
Types are composed of type variable $\tvar$, parameterized type constructors
$\T{t}$ and arrow types annotated with a kind, noted $\tau \tarr{k} \tau$.
Kinds can be either a kind variable $\kvar$, affine $\kaff$ or unrestricted $\kun$.
Type schemes and kind schemes are guarded by a set of constraints.

@@

\begin{figure*}[t]
  \centering
  \input{grammar}
  \caption{Syntax}
  \label{grammar}
\end{figure*}
\input{internal-grammar}

\clearpage
\subsection{Semantics}
\input{fig-reduction}

\clearpage
\subsection{Typing}

\subsubsection{Constraint language}

\newcommand\A{\mathcal A}
\newcommand\SC{\mathcal S}

Let us note $\A$ our constraint system. The grammar of constraints is
given in  \cref{grammar:constraint}.
$\A$ is defined as the smallest cylindric term constraint system that
satisfies the axiom shown in \cref{rules:entail}.
We follows the traditional HM(X) formulation
with conjunctions, projections and type inequalities.
The new element specific to our approach are kind inequalities.
Entailment is noted $\entail{C}{D}$, where $D$ is a consequence of the
constraints $C$.
We say that $C$ and $D$ are equivalent, noted $C \equivC D$,
when $\entail{C}{D}$ and $\entail{D}{C}$.
\TODO{Give the cylindric properties ?}

\begin{figure}[tp]
  \centering
  \begin{align*}
    C &::= \Cleq{\tau_1}{\tau_2}
        \mid \Cleq{k_1}{k_2}
        \mid C_1 \Cand C_2
        \mid \Cproj{\tvar}{C}
        \mid \Cproj{\kvar}{C}
  \end{align*}
  \caption{The constraint language}
  \label{grammar:constraint}
\end{figure}

\begin{figure*}[tp]
  \input{infer/lattice}
  \caption{Lattice inequalities -- $k \lk_\Lat k'$}
  \input{infer/entails}
  \caption{Base entailment rules -- $\entail{C}{D}$ }
  \label{rules:entail}
\end{figure*}

We note $\SC$ the set of solved forms
which can be used inside type and kind schemes.
We define $\SC$ as $\A$ quotiented by the relation $\equivC$.
%
We consider the existence of a function $\operatorname{normalize}$ which takes
a constraint in $\A$ and a substitution $\psi$ and returns a constraint
in solved form $C' \in \SC$,
and an updated substitution. We detail the implementation
of the normalization function in \cref{sec:normalize}

% $\mathcal S$ is composed only of kind
% inequalities \emph{over variables}. For convenience, if $C\in\mathcal S$, we
% note $C$ as a list of kind inequalities: $\Cleq{\kvar_i}{\kvar_{i'}}^n$.
% \TODO{Extend the properties of solved forms}

\subsubsection{Kinding}

\begin{figure}[ht]
  \centering
  \input{infer/sd-kinds}
  \caption{Syntax-directed kind inference --
    $\inferSK{C}{\E}{\tau}{(\Multi k)k}$}
  \label{rules:sd-kinding}
\end{figure}

\subsubsection{Environments}

\begin{figure}[tp]
  \input{infer/sd-splitting}
  \caption{Splitting --- environments $\lsplit
    C\E\E\E$; binders $\bsplit Cbbb$}
  \label{fig:sd-splitting}
\end{figure}

\begin{figure}[tp]
  \input{infer/sd-borrowing}
  \caption{Borrowing --- environments $\lsplit
    C\E\E\E$; binders $\bsplit Cbbb$}
  \label{fig:sd-borrowing}
\end{figure}

\subsubsection{Typing}

\begin{figure*}[htp]
  \input{infer/sd-typing}
  \caption{Syntax-directed typing rules}
  \label{fig:syntax-directed-typing}
\end{figure*}

\clearpage
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:
